The Atala PRISM Protocol

### 2022-11-07 prism-backend/research/prism-protocol/did_atala.tex
1st potential protocol for Atala DID.

    \begin{minipage}[t]{0.55\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Append} \land \arg = x$}{
        \pccomment{\fbox{Append data to \aobbl}} \\
        \aobbl \gets \aobbl || x \\
        \pcif P~\text{is corrupt}:~\ucio{Send}~(\uccmd{Appended},x,P)~\text{to}
        ~\adv
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.45\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Retrieve} \land \arg =
        \emptyset$}{
        \pccomment{\fbox{Retrieve the contents of \aobbl}} \\
        \text{Output}~(\uccmd{Retrieved},\aobbl)~\text{to}~P
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal append-only bulletin board functionality, \IdealFaobb, from
    \cite{acc+20}.}
  \label{fig:faobb}
\end{figure}

@@

    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{KeyGen} \land \arg =
        \emptyset$}{
        \pccomment{\fbox{Key Generation}} \\
        \pcif \sid = (P,\ssid): \\
        \pcind \ucio{Send}~(\uccmd{KeyGen},\sid)~\text{to}~\adv \\
        \pcind \ucio{Receive}~(\uccmd{VerKey},\sid,v)~\text{from}~\adv \\
        \pcind \ucio{Output}~(\uccmd{VerKey},\sid,v)~\text{to}~P \\
        \pcind \text{Store}~(P,v)
      }
      \vspace*{0.25em}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Verify} \land
        \arg = (\sig,\msg,v')$}{
        \pccomment{\fbox{Verify a message}} \\
        \ucio{Send}~(\uccmd{Verify},\sid,\msg,\sig,v)~\text{to}~\adv \\
        \ucio{Receive}~(\uccmd{Verified},\sid,\msg,\phi)~\text{from}~\adv \\
        \pcif v' = v \land (\msg,\sig,v,1) \in \SIG: f \gets 1 \\
        \pcelse \pcif v' = v \land P~\text{is not corrupt}~\land \\
        \pcind (\msg,\sig',v,1) \notin \SIG: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v,0) \rbrace;
        f \gets 0 \\
        \pcelse \pcif (\msg,\sig,v',f') \in \SIG: f \gets f' \\
        \pcelse: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v',\phi) \rbrace;~
        f \gets \phi \\
        \ucio{Output}~(\uccmd{Verified},\sid,\msg,f)
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Sign} \land \arg = \msg$}{
        \pccomment{\fbox{Sign a message}} \\
        \pcif \sid = (P,\ssid) \\
        \pcind \ucio{Send}~(\uccmd{Sign},\sid,\msg)~\text{to}~\adv \\
        \pcind \ucio{Receive}~(\uccmd{Signature},\sid,\msg,\sig)~\text{from}~
        \adv \\
        \pcind \pcif (\msg,\sig,v,0) \in \SIG: \\
        \pcind \pcind \ucio{Output}~(\uccmd{Error}) \\
        \pcind \pcelse: \\
        \pcind \pcind \SIG \gets \SIG \cup (\msg,\sig,v,1) \\
        \pcind \pcind \ucio{Output}~(\uccmd{Signature},\sid,\msg,\sig)
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal signature functionality, \IdealFSig, from \cite{canetti03}.}
  \label{fig:faobb}
\end{figure}

@@

\begin{figure}
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealFPKIDID^{\P}(\cmd,\sid,\arg)$
        from some party $P$ to run command \cmd, in session \sid, with arguments
        \arg, $\IdealFPKIDID^{\P}$ operates as follows.}
      \textrm{If the received \cmd and \arg pair does not meet any of
        the following options, the request is ignored.}
      \textrm{\P is a boolean predicate defining concrete conditions for
        acceptable DID creations, updates and deactivations.
      }
      \textrm{\todo{We assume that the functionality keeps track of the
          received requests.}}
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.55\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
        \lbrace (i,\lbl_i) \rbrace_{i\in[n]}$}{
        \pccomment{\fbox{DID Create operation}} \\
        \ucio{Send}~(\uccmd{Create}, \sid, \arg)~\text{to}~\adv \\
        \ucio{Receive}~(\uccmd{CreateOk}, \sid, did,
        \lbrace \val_i \rbrace_{i\in[n]}) \\
        \pcind \text{from}~\adv \\
        \pcif (P,\did,\cdot) \notin \DID \land
        \P(\lbrace \lbl_i \rbrace_{i\in[n]}) = 1: \\
        \pcind \DID \gets \DID \cup \lbrace (P,did,\lbrace
        (\lbl_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
        \ucio{Output}~(\uccmd{Created},\sid,\did,
        \lbrace \val_i \rbrace_{i\in[n]})
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
        \arg = did$}{
        \pccomment{\fbox{DID Read operation}} \\
        \ucio{Send}~(\uccmd{Read},\sid,\arg)~\text{to}~\adv \\
        \ucio{Receive}~(\uccmd{ReadOk},\sid)~\text{from}~\adv \\
        \pcif (\cdot,did,\sval) \in \DID: \\
        \pcind \ucio{Output}~(\uccmd{Read},\sid,(did,\sval))~\text{to}~P \\
        \pcelse: \\
        \pcind \ucio{Output}~(\uccmd{Read},\sid,\bot)~\text{to}~P    
      }    
    \end{minipage}
    \vspace*{0.25em} \\
    \begin{minipage}[t]{0.55\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
        \arg = (did, \lbrace (i,\lbl_i) \rbrace_{i\in[n]})$}{      
        \pccomment{\fbox{DID Update operation}} \\
        \ucio{Send}~(\uccmd{Update}, \sid, \arg)~\text{to}~\adv \\
        \ucio{Receive}~(\uccmd{UpdateOk},\sid,
        \lbrace \val_i \rbrace_{i\in[n]}) \\
        \pcind \text{from}~\adv \\
        \pcif (P,did,\sval) \in \DID \land
        \P(\lbrace \lbl_i \rbrace_{i\in[n]}) = 1: \\
        \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval) \rbrace~\cup \\
        \hspace*{7em} \lbrace (P, did,
        \lbrace \lbl_i,\val_i \rbrace_{i\in[n]}) \rbrace
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
        (did, \lbrace (i,\lbl_i) \rbrace_{i\in[n]})$}{
        \pccomment{\fbox{DID Deactivate operation}} \\
        \ucio{Send}~(\uccmd{Deact}, \sid, \arg)~\text{to}~\adv \\
        \ucio{Receive}~(\uccmd{DeactOk},\sid)~\text{from}~\adv \\
        \pcif (P,did,\sval) \in \DID~\land \\
        \pcind \lbrace \lbl_i \rbrace_{i\in[n]} \subseteq \LBL(\sval): \\
        \pcind \sval' \gets \sval \setminus \lbrace (\lbl_i,\cdot)
        \rbrace_{i\in[n]} \\
        \pcind \pcif \P(\sval') = 1: \\
        \pcind \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval) \rbrace~\cup \\
        \hspace*{8em}\lbrace (P,did,\sval') \rbrace
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealFPKIDID.}
  \label{fig:fpkidid}
\end{figure}

@@

\paragraph{Key Types.} %
All operations in Atala PRISM are authenticated, meaning that a cryptographic
key needs to be involved in proving that the operation comes from the intended
party. Currently, 4 types of keys are supported: master keys (\MasterKey), for
operating on DIDs; issuing keys, for issuing VCs (\IssueKey); communication
keys (\CommKey), for key-exchange protocols; and authentication keys, to
associate to VCs (\AuthKey).

\paragraph{DID Operations.} %
\todo{For now, this ignores \emph{unpublished} DIDs.}
\figref{fig:alapkidid} contains the specification of Atala PRISM's DID method,
which we describe next.

\begin{description}
\item{\uccmd{Create}.} %
  To create a DID, the user has to specify the key pairs to create, along with
  fresh identifiers and their type. For ease of exposition, we assume that the
  used identifiers just follow a sequence. The predicate \P in Atala simply
  requires that at least one key pair of \MasterKey type is created. The
  \uccmd{Create} algorithm creates as many keys of each type as requested,
  and keeps a list with the identifier, type and public key of each key pair.
  This list is then hashed into a value \texttt{h}, and the created DID is
  ``\texttt{\did:prism:h}''. A transaction containing the created list, the
  DID, and the \texttt{CreateDID} instruction is sent to the blockchain,
  signed by one of the master keys -- w.l.o.g., we simply use the first one
  in the list. The DID owner stores locally the list, along with the
  corresponding private keys.
\item[\uccmd{Read}.] %
  Getting the contents of a DID simply requires querying a \todo{PRISM node} for
  it. The PRISM node returns the associated list, along with all related
  \texttt{CreateDID} and \texttt{UpdateDID} transactions in the blockchain. If
  the signatures of these transactions are valid, and the result of applying in
  order all the operations produces the list returned by the \todo{PRISM node},
  then the list is an up to date DID Document of the corresponding DID.
\item[\uccmd{Update}.] %
  Updating a DID requires to (optionally) specify a list of previously existing
  key identifiers to remove, and (optionally) specify a list of new key
  identifiers, and their types, to add. To simplify notation, we just adopt the
  (equivalent) convention that the caller indicates a list of $(i,\lbl_i)$
  pairs, which is interpreted as follows. If $i$ is an already existing
  identifier, the label is ignored and the corresponding key is not removed. If
  $i$ is a new identifier, then a new key pair of the given type is created.
  Omitting an existing identifier is then interpreted as the desire to remove
  the corresponding key pair. Otherwise, the same policy as for \uccmd{Create}
  applies here. Key pairs for the new keys are generated, and a transaction is
  sent to the blockchain, signed with one of the master keys in the DID
  \emph{before applying the update}. This transaction contains a list of the
  identifiers of the keys to remove, and a list of the identifiers and types
  of the keys to add.
\item[\uccmd{Deactivate}.] %
  A deactivate operation is simply an update operation with no new key
  identifiers, and that only specifies the existing identifiers that will
  \emph{not} be deactivated.
\end{description}

\begin{figure}
  \begin{framed}
    \begin{minipage}[t]{\textwidth}
      \textrm{Upon receiving a call $\IdealFPKIDID^{\P}(\cmd,\sid,\arg)$
        from some party $P$ to run command \cmd, in session \sid, with arguments
        \arg, $\IdealFPKIDID^{\P}$ operates as follows.}
      \textrm{If the received \cmd and \arg pair does not meet any of
        the following options, the request is ignored.}
      \textrm{\P is a boolean predicate defining concrete conditions for
        acceptable DID creations, updates and deactivations.
      }
      \textrm{\todo{We assume that the functionality keeps track of the
          received requests.}}
    \end{minipage}
    \vspace*{0.5em}

    \begin{minipage}[t]{0.55\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
        \lbrace (i,\lbl_i) \rbrace_{i\in[n]}$}{
        \pccomment{\fbox{DID Create operation}} \\
        \st \gets \emptyset \\
        \pcif \exists i~\suchthat~\lbl_i \notin [\MasterKey,\AuthKey,
        \CommKey,\IssueKey] \lor
        \nexists i~\suchthat~\lbl_i = \MasterKey: \\
        \pcind \texttt{Abort} \\
        \pcfor i \in [n]: \\
        \pcind \sid' \gets (P,\sid,i) \\
        \pcind \ucio{Send}~(\uccmd{KeyGen},\sid')~\text{to}~\IdealFSig \\
        \pcind \ucio{Receive}~(\uccmd{VerKey},\sid',\pk_i)~\text{from}~
        \IdealFSig \\
        \pcind \st \gets \st~||~(i,\lbl_i,\pk_i) \\
        \ucio{Send}~(\uccmd{Hash},\st)~\text{to}~\IdealFRO \\
        \ucio{Receive}~(\uccmd{Hashed},\st,h)~\text{from}~\IdealFRO \\
        \did \gets ``did:prism:''||~h \\
        \text{Locally store}~(\did,\lbrace (i,\lbl_i,\pk_i) \rbrace_{i\in[n]}) \\
        \tx \gets (\texttt{CreateDID}, \did, \st) \\
        \text{Let}~j\in[n]~\text{be the least value s.t.}~\lbl_j = \MasterKey \\
        \ucio{Send}~(\uccmd{Sign},(P,\sid,j),\msg)~\text{to}~\IdealFSig \\
        \ucio{Receive}~(\uccmd{Signature},(P,\sid,j),\msg,\sig)~\text{from}~\IdealFSig \\
        \ucio{Send}~(\uccmd{Append},(\msg,\sig))~\text{to}~\IdealFaobb \\
        \ucio{Output}~(\uccmd{Created},\sid,\did,\lbrace \pk_i \rbrace_{i\in[n]})
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
        \arg = did$}{
        \pccomment{\fbox{DID Read operation}} \\  
      }    
    \end{minipage}
    \vspace*{0.25em} \\
    \begin{minipage}[t]{0.55\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
        \arg = (did, \lbrace (i,\lbl_i) \rbrace_{i\in\tilde{N}\subseteq [\tilde{n}]})$}{
        \pccomment{\fbox{DID Update operation}} \\
        \pcif \exists i~\suchthat~\lbl_i \notin [\MasterKey,\AuthKey,
        \CommKey,\IssueKey] \lor
        \nexists i~\suchthat~\lbl_i = \MasterKey: \\
        \pcind \texttt{Abort} \\
        \text{Read}~(\did,\lbrace (i,\l_i,\pk_i) \rbrace_{i\in[n]})~
        \text{from local storage} \\
        \st_{add} \gets \emptyset \\
        \pcfor i > n \in [\tilde{N}]: \\
        \pcind \ucio{Send}~(\uccmd{KeyGen},(P,\sid,i))~\text{to}~\IdealFSig \\
        \pcind \ucio{Receive}~(\uccmd{VerKey},(P,\sid,i),\pk_i)~\text{from}~
        \IdealFSig \\
        \pcind \st_{add} \gets \st_{add}~||~(i,\lbl_i,\pk_i) \\
        \st_{del} \gets [n] \setminus [\tilde{N}] \\
        \text{Locally store}~(\did,\lbrace (i,\lbl_i,\pk_i) \rbrace_{i\in[\tilde{n}]} \\
        \tx \gets \texttt{UpdateDID}~||~\did~||~\st_{del}~||~\st_{add} \\
        \text{Let}~j\in[n]~\text{be the least value s.t.}~\lbl_j = \MasterKey \\
        \ucio{Send}~(\uccmd{Sign},(P,\sid,j),\msg)~\text{to}~\IdealFSig \\
        \ucio{Receive}~(\uccmd{Signature},(P,\sid,j),\msg,\sig)~\text{from}~\IdealFSig \\
        \ucio{Send}~(\uccmd{Append},(\msg,\sig))~\text{to}~\IdealFaobb \\
        \ucio{Output}~(\uccmd{Updated},\sid,\did,\lbrace \pk_i \rbrace_{i\in[\tilde{N}]})
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
        (did, \lbrace (i,\lbl_i) \rbrace_{i\in[n]})$}{
        \pccomment{\fbox{DID Deactivate operation}} \\
      }
    \end{minipage}
  \end{framed}
  \caption{Atala PRISM DID-based PKI construction, \RealPKIDIDAtala.}
  \label{fig:fpkidid}
\end{figure}

% \begin{figure}

>>>>

    \begin{minipage}[t]{0.55\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Append} \land \arg = x$}{
        \pccomment{\fbox{Append data to \aobbl}} \\
        \aobbl \gets \aobbl || x \\
        \pcif P~\text{is corrupt}:~\ucsend~(\uccmd{Appended},x,P)~\text{to}
        ~\adv
      }
    \end{minipage}
    \hspace*{1.25em}
    \begin{minipage}[t]{0.45\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Retrieve} \land \arg =
        \emptyset$}{
        \pccomment{\fbox{Retrieve the contents of \aobbl}} \\
        \text{Output}~(\uccmd{Retrieved},\aobbl)~\text{to}~P
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal append-only bulletin board functionality, \IdealFaobb, from
    \cite{acc+20}.}
  \label{fig:faobb}
\end{figure}

@@

    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{KeyGen} \land \arg =
        \emptyset$}{
        \pccomment{\fbox{Key Generation}} \\
        \pcif \sid = (P,\ssid): \\
        \pcind \ucsend~(\uccmd{KeyGen},\sid)~\text{to}~\adv \\
        \pcind \ucrecv~(\uccmd{VerKey},\sid,v)~\text{from}~\adv \\
        \pcind \ucio{Output}~(\uccmd{VerKey},\sid,v)~\text{to}~P \\
        \pcind \text{Store}~(P,v)
      }
      \vspace*{0.25em}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Verify} \land
        \arg = (\sig,\msg,v')$}{
        \pccomment{\fbox{Verify a message}} \\
        \ucsend~(\uccmd{Verify},\sid,\msg,\sig,v)~\text{to}~\adv \\
        \ucrecv~(\uccmd{Verified},\sid,\msg,\phi)~\text{from}~\adv \\
        \pcif v' = v \land (\msg,\sig,v,1) \in \SIG: f \gets 1 \\
        \pcelse \pcif v' = v \land P~\text{is not corrupt}~\land \\
        \pcind (\msg,\sig',v,1) \notin \SIG: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v,0) \rbrace;
        f \gets 0 \\
        \pcelse \pcif (\msg,\sig,v',f') \in \SIG: f \gets f' \\
        \pcelse: \\
        \pcind \SIG \gets \SIG \cup \lbrace (\msg,\sig,v',\phi) \rbrace;~
        f \gets \phi \\
        \ucio{Output}~(\uccmd{Verified},\sid,\msg,f)
      }
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
      \procedure[linenumbering]{$\pcif \cmd = \uccmd{Sign} \land \arg = \msg$}{
        \pccomment{\fbox{Sign a message}} \\
        \pcif \sid = (P,\ssid) \\
        \pcind \ucsend~(\uccmd{Sign},\sid,\msg)~\text{to}~\adv \\
        \pcind \ucrecv~(\uccmd{Sig.},\sid,\msg,\sig)~\text{from}~
        \adv \\
        \pcind \pcif (\msg,\sig,v,0) \in \SIG: \\
        \pcind \pcind \ucio{Output}~(\uccmd{Error}) \\
        \pcind \pcelse: \\
        \pcind \pcind \SIG \gets \SIG \cup (\msg,\sig,v,1) \\
        \pcind \pcind \ucio{Output}~(\uccmd{Sig.},\sid,\msg,\sig)
      }
    \end{minipage}
  \end{framed}
  \caption{Ideal signature functionality, \IdealFSig, from \cite{canetti03}.}
  \label{fig:faobb}
\end{figure}

@@

\begin{figure}
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Upon receiving a call $\IdealFPKIDID^{\P}(\cmd,\sid,\arg)$
          from some party $P$ to run command \cmd, in session \sid, with arguments
          \arg, $\IdealFPKIDID^{\P}$ operates as follows.}
        \textrm{If the received \cmd and \arg pair does not meet any of
          the following options, the request is ignored.}
        \textrm{\P is a boolean predicate defining concrete conditions for
          acceptable DID creations, updates and deactivations.
        }
        \textrm{\todo{We assume that the functionality keeps track of the
            received requests.}}
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (i,\lbl_i) \rbrace_{i\in[n]}$}{
          \pccomment{\fbox{DID Create operation}} \\
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{CreateOk}, \sid, did,
          \lbrace (i, \val_i) \rbrace_{i\in[n]}) \\
          \pcind \text{from}~\adv \\
          \pcif (P,\did,\cdot) \notin \DID \land
          \P(\lbrace (i, \lbl_i) \rbrace_{i\in[n]}) = 1: \\
          \pcind \DID \gets \DID \cup \lbrace (P,did,\lbrace
          (i,\lbl_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          \ucio{Output}~(\uccmd{Created},\sid,\did,
          \lbrace (i,\val_i) \rbrace_{i\in[n]})
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \pccomment{\fbox{DID Read operation}} \\
          \ucsend~(\uccmd{Read},\sid,\arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{ReadOk},\sid)~\text{from}~\adv \\
          \pcif (\cdot,did,\sval) \in \DID: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,(did,\sval))~\text{to}~P \\
          \pcelse: \\
          \pcind \ucio{Output}~(\uccmd{Read},\sid,\bot)~\text{to}~P    
        }    
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \lbrace (i,\lbl_i) \rbrace_{i\in[n]})$}{      
          \pccomment{\fbox{DID Update operation}} \\
          \ucsend~(\uccmd{Update}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{UpdateOk},\sid,
          \lbrace (i,\val_i) \rbrace_{i\in[n]}) \\
          \pcind \text{from}~\adv \\
          \pcif (P,did,\sval) \in \DID \land
          \P(\lbrace (i,\lbl_i) \rbrace_{i\in[n]}) = 1: \\
          \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval) \rbrace~\cup \\
          \hspace*{7em} \lbrace (P, did,
          \lbrace (i,\lbl_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          \ucio{Output}~(\uccmd{Updated},\sid,
          (\did,\lbrace (i,\val_i) \rbrace_{i\in[n]}))
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace (i,\lbl_i) \rbrace_{i\in[n]})$}{
          \pccomment{\fbox{DID Deactivate operation}} \\
          \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\adv \\
          \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\adv \\
          \pcif (P,did,\sval) \in \DID~\land \\
          \pcind \lbrace (i,\lbl_i) \rbrace_{i\in[n]} \subseteq \LBL(\sval): \\
          \pcind \sval' \gets \sval \setminus \lbrace (i,\lbl_i,\cdot)
          \rbrace_{i\in[n]} \\
          \pcind \pcif \P(\sval') = 1: \\
          \pcind \pcind \DID \gets \DID \setminus \lbrace (P,did,\sval) \rbrace~\cup \\
          \hspace*{8em}\lbrace (P,did,\sval') \rbrace \\
          \ucio{Output}~(\uccmd{Deacted},\sid,(\did,\sval'))
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealFPKIDID.}
  \label{fig:fpkidid}
\end{figure}

@@

\paragraph{Key Types.} %
All operations in Atala PRISM are authenticated, meaning that a cryptographic
key needs to be involved in proving that the operation comes from the intended
party. Atala specifies four key types, that are thus supported in Atala's DID
method. Namely: master keys (\MasterKey), for operating on DIDs; issuing keys,
for issuing VCs (\IssueKey); communication keys (\CommKey), for key-exchange
protocols; and authentication keys, to associate to VCs (\AuthKey).

\paragraph{Custom Notation.} %
Keys in Atala are identified by indexes. In the following description of the
\RealPKIDIDAtala construction, given a set of keys, we frequently refer to the
``master key with lowest index''.  For brevity, given a set \st, we use
$\LMKI(\st)$ to denote such index.

\paragraph{DID Operations.} %
\todo{For now, this ignores \emph{unpublished} DIDs.}
\figref{fig:atalapkidid1} and \figref{fig:atalapkidid2} contain the
specification of Atala PRISM's DID method,
which we describe next. Looking at the \texttt{Create} and \texttt{Update}
operations, it is clear that $\P(\lbrace \lbl_i \rbrace_{i\in[n]})$ in Atala
requires that all key labels are of a known type (\MasterKey, \IssueKey,
\CommKey, or \AuthKey) and that at least one of the keys in the DID Document
is of type \MasterKey.

\begin{description}
\item{\uccmd{Create}.} %
  To create a DID, the user has to specify the key pairs to create, along with
  fresh identifiers and their type. For ease of exposition, we assume that the
  used identifiers just follow a sequence. The predicate \P in Atala simply
  requires that at least one key pair of \MasterKey type is created. The
  \uccmd{Create} algorithm creates as many keys of each type as requested,
  and keeps a list with the identifier, type and public key of each key pair.
  This list is then hashed into a value \texttt{h}, and the created DID is
  ``\texttt{\did:prism:h}''. A transaction containing the created list, the
  DID, and the \texttt{CreateDID} instruction is sent to the blockchain,
  signed by one of the master keys -- w.l.o.g., we simply use the first one
  in the list. The DID owner stores locally the list, along with the
  corresponding private keys.
\item[\uccmd{Read}.] %
  Getting the contents of a DID simply requires processing the associated
  transactions in the blockchain. Namely, transactions $(\tx,\sig)$ where $\tx$
  is either $(\texttt{CreateDID},\did,\cdot)$ or $(\texttt{UpdateDID},\did,
  \cdot)$. The associated DID Document is the result of applying in an ordered
  manner (from oldest to newest) all such validly signed transactions.
\item[\uccmd{Update}.] %
  Updating a DID requires to (optionally) specify a list of previously existing
  key identifiers to remove, and (optionally) specify a list of new key
  identifiers, and their types, to add. To simplify notation, we just adopt the
  (equivalent) convention that the caller indicates a list of $(i,\lbl_i)$
  pairs, which is interpreted as follows. If $i$ is an already existing
  identifier, the label is ignored and the corresponding key is not removed. If
  $i$ is a new identifier, then a new key pair of the given type is created.
  Omitting an existing identifier is then interpreted as the desire to remove
  the corresponding key pair. Otherwise, the same policy as for \uccmd{Create}
  applies here. Key pairs for the new keys are generated, and a transaction is
  sent to the blockchain, signed with one of the master keys in the DID
  \emph{before applying the update}. This transaction contains a list of the
  identifiers of the keys to remove, and a list of the identifiers and types
  of the keys to add.
\item[\uccmd{Deactivate}.] %
  A deactivate operation is simply an update operation with no new key
  identifiers, and that only specifies the existing identifiers that will
  \emph{not} be deactivated.
\end{description}

\begin{figure}[ht!]
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \lbrace (i,\lbl_i) \rbrace_{i\in[n]}$}{
          \pcif \exists i~\suchthat~\lbl_i \notin [\MasterKey,\AuthKey,
          \CommKey,\IssueKey] \lor
          \nexists i~\suchthat~\lbl_i = \MasterKey: \\
          \pcind \texttt{abort} \\
          \st \gets \emptyset \\
          \pcfor i \in [n]: \\
          \pcind \sid' \gets (P,\sid,i) \\
          \pcind \ucsend~(\uccmd{KeyGen},\sid')~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{VerKey},\sid',\pk_i)~\text{from}~
          \IdealFSig \\
          \pcind \st \gets \st \cup \lbrace (i,\lbl_i,\pk_i) \rbrace \\
          \ucsend~(\uccmd{Hash},\st)~\text{to}~\IdealFRO \\
          \ucrecv~(\uccmd{Hashed},\st,h)~\text{from}~\IdealFRO \\
          \did \gets ``did:prism:''||~h \\
          \tx \gets (\texttt{CreateDID}, \did, \st) \\
          \text{Let}~j\in[n]~\text{be the least value s.t.}~\lbl_j = \MasterKey \\
          \ucsend~(\uccmd{Sign},(P,\sid,j),\tx)~\text{to}~\IdealFSig \\
          \ucrecv~(\uccmd{Sig.},(P,\sid,j),\tx,\sig)~\text{from}~\IdealFSig \\
          \ucsend~(\uccmd{Append},(\msg,\sig))~\text{to}~\IdealFaobb \\
          \ucio{Output}~(\uccmd{Created},\sid,\did,\lbrace \pk_i \rbrace_{i\in[n]})
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering=on]{$\pcif \cmd = \uccmd{Read} \land
          \arg = did$}{
          \ucsend~(\uccmd{Retrieve})~\texttt{to}~\IdealFaobb \\
          \ucrecv~(\uccmd{Retrieve},L)~\texttt{from}~\IdealFaobb \\
          \st \gets \emptyset \\
          L_{\did} \gets (\ast,\did,\ast) \subseteq \DID \\
          \pcfor (\tx=(\texttt{op},\did,\cdot),\sig) \in L_{\did}: \\
          \pcind \pk \gets \emptyset \\
          \pcind \pcif \texttt{op} = \texttt{CreateDID}~\land \\
          \pcind \pcind \text{no \texttt{CreateDID} appeared before}: \\
          \pcind \pcind (\cdot,\cdot,\lbrace (i,\lbl_i,\pk_i)
          \rbrace_{i\in[n]}) \gets \tx \\
          \pcind \pcind \st \gets \lbrace (i,\lbl_i,\pk_i) \rbrace_{i \in [n]} \\
          \pcind \pcind j \gets \LMKI(\st); \pk \gets \pk_j \\
          \pcind \pcif \texttt{op} = \texttt{UpdateDID}: \\
          \pcind \pcind (\cdot,\cdot,(\st_{del},\st_{add})) \gets \tx \\
          \pcind \pcind \lbrace i \rbrace_{i\in[n]} \gets \st_{del} \\ 
          \pcind \pcind \lbrace (i,\lbl_i,\pk_i) \rbrace_{i\in[\tilde{n}]} \gets \st_{add} \\
          \pcind \pcind j \gets \LMKI(\st); \pk \gets \pk_j \\
          \pcind \pcind \st \gets \st \setminus \st_{del} \cup \st_{add} \\
          \pcind \pcif \pk = \emptyset: \text{abort} \\
          \pcind \ucsend~(\uccmd{Verify},(P,\sid),\tx,\sig,\pk)~\text{to}~\IdealFSig \\
          \pcind \ucrecv~(\uccmd{Verified},(P,\sid),\tx,b)~\text{from}~\IdealFSig \\
          \pcind \pcif b = 0: \text{abort} \\
          \ucio{Output}~(\uccmd{Read},\sid,(\did,\st))
        }    
      \end{minipage}
    }
  \end{framed}
  \caption{\texttt{Create} and \texttt{Read} operations in Atala
    PRISM DID-based PKI protocol, \RealPKIDIDAtala, realising \IdealFPKIDID
    in the $(\IdealFRO,\IdealFSig,\IdealFaobb)$-hybrid model.}
  \label{fig:atalapkidid1}
\end{figure}

\begin{figure}[ht!]
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{We assume that the party running
          \texttt{Update} and \texttt{Deactivate} operations has an up to date
          copy $(\did,\lbrace (i,\lbl_i,\pk_i) \rbrace_{i\in[n]})$ of the DID to
          modify. Otherwise, a \texttt{Read} operation needs to be run before.}
      \end{minipage}
    }
    \vspace*{0.5em}
